#===================================================================
# ISM645 / IAF601   Principle of Predictive Analytics
# Final Exam        Predictive Modeling
# Due Date          December 1, 11:59 pm
#Vathana Him
#===================================================================


#===================================================================
############################ Final Exam ############################
#===================================================================
```{r}
library(tidyverse)
library(ggplot2)
library(rsample)
library(DescTools)
library(caTools)
library(caret)
library(broom)
library(yardstick)
library(randomForest)
library(cutpointr)
```

# Import the student_performance.csv and explore it.
#====================== Write R code HERE ==========================


```{r}
data = read.csv('student_performance.csv')
head(data)
summary(data)
```



#===================================================================



#====================================
#======= Question 1 (1 Point) =======
# Answer the following questions by presenting a relevant plot (visualization).
# Q1-1. Do grades of Mathematics and Portuguese differ by gender?
# Q1-2. Does study time matter to get better grades of Mathematics and Portuguese?
# Q1-3. Does access to Internet influence students' grades of Mathematics and Portuguese?

## There could be various approaches to answer to these questions.

#====================== Write R code HERE ==========================

```{r}
#Q1-1
ggplot(data) + geom_bar(aes(subject, total_grade, fill = sex), position = 'dodge', stat='summary', fun='mean')+labs(title='Total_grade By Gender and Subject')

#Q1-2
ggplot(data, aes(x=studytime, y=total_grade)) + geom_point(size=3, aes(color=subject))+labs(title='Study time v. Total Grade By Subject')

#Q1-3
ggplot(data)+geom_bar(aes(internet, total_grade, fill = subject), position = 'dodge', stat='summary', fun='mean')+labs(title='Internet Access by Grade and Subject')

```






#===================================================================



#====================================
#======= Question 2 (2 Point) =======
# Q2-1. Split the data into two separate data frames for each subject: Mathematics and Portuguese. Then, remove a variable "subject".
# Q2-2. Build linear regression models to predict total_grade of Mathematics and Portuguese, respectively, based on time spending patterns (travel time, study time, free time after school, and going out with friends)
# Q2-3. Based on your model, make prediction for a new student's final grades of Mathematics and Portuguese, based on the following information.
#       You can write your prediction results by using comment (#).

#------------------ New Student's Information
#------------------ traveltime = 2
#------------------ studytime = 4
#------------------ freetime = 3
#------------------ goout = 2

#====================== Write R code HERE ==========================

```{r}
#Q2-1
math_data = data %>% filter(subject == 'Mathematics') %>% select(-c(subject))
portuguese_data = data %>% filter(subject == 'Portuguese') %>% select(-c(subject))

#Q2-2
math_reg = lm(total_grade~traveltime+studytime+freetime+goout, data=math_data)
portuguese_reg = lm(total_grade~traveltime+studytime+freetime+goout, data=portuguese_data)

#Q2-3
new_student = data.frame(traveltime=c(2), studytime=c(4), freetime=c(3), goout=c(2))

pred_math_grade = math_reg %>% predict(newdata = new_student)
pred_math_grade

pred_portuguese_grade = portuguese_reg %>% predict(newdata = new_student)
pred_portuguese_grade

#The predicted total grade for math grade with the new student parameter is ~35.9 and the predicted total grade for #portuguese grade with the new student parameter is ~39.8.

```







#===================================================================



#### For the subsequent questions (Q3 to Q5), use the dataset for grades of Mathematics.

#====================================
#======= Question 3 (2 Point) =======

# Q3-1. Create a new binary variable, "pass", defined as TRUE if total_grade > 30 and FALSE otherwise.
#       To avoid any error, remove the variable, "subject".
# Q3-2. Split data into 70% train and 30% test data.
# Q3-3. Based on train data, build a logistic regression model to predict "pass" using all available predictor variables.
#       Note that "total_grade" should not be included as a predictor.
# Q3-4. Based on test data, draw a ROC curve and calculate AUC for the model trained in Q3-3.

#====================== Write R code HERE ==========================

```{r}
#Q3-1
math_data = math_data %>% mutate(pass = if_else(total_grade>30, 1, 0))

#Q3-2
math_split = initial_split(math_data, 0.7)
math_train = training(math_split)
math_test = testing(math_split)

#Q3-3
log_reg = glm(pass~.-total_grade, data=math_train, family='binomial')

#Q3-4
pred_log_reg = log_reg %>% augment(newdata = math_test, type.predict = 'response')
roc = roc(pred_log_reg, x=.fitted, class=pass, pos_class=1, neg_class=0, direction = ">=")
plot(roc)
auc(roc)
```









#===================================================================



#====================================
#======= Question 4 (2 Point) =======
# Q4-1. Based on train data, build a random forest model to predict "pass" using all available predictor variables.
# Q4-2. Based on test data, draw a ROC curve and calculate AUC for the model trained in Q4-1.
# Q4-3. According to the results of Q3 and Q4, which model do you think is better? Why?
#       Write your answer below by using comment (#).

#====================== Write R code HERE ==========================

```{r}
#Q4-1
math_rf = randomForest(as.factor(pass)~.-total_grade, data = math_train, ntree=1200, importance=TRUE)

#Q4-2
math_pred = math_rf %>% predict(newdata= math_test, type="prob")
math_test = math_test %>% mutate(.fitted_rf = math_pred[, 2])
roc = roc(math_test, x=.fitted_rf, class=pass, pos_class=1, neg_class=0)
plot(roc)
auc(roc)

#Q4-3

#According to the results of Q3 and Q4, the random forest model of Q4 is a better fit because the AUC score for Q4 is #higher than that of Q3 with a score of 0.67 and 0.64 respectively. 
```









#===================================================================


#====================================
#======= Question 5 (2 Point) =======
# For this question, use the random forest model as in Q4.

# Q5. One might argue that background of parents and family structure are important in predicting children's math performance. 
#     Build another random forest model with only family-related factors (Pstatus, Medu, Mjob, Fedu, Fjob, guardian, famsize, famrel, and famsup).
#     Based on your analysis, do you agree or disagree with this argument?
#     Write your answer below by using comment (#).

#====================== Write R code HERE ==========================


```{r}
#Q5
q5_math_rf = randomForest(as.factor(pass)~Pstatus+Medu+Mjob+Fedu+Fjob+guardian+famsize+famrel+famsup, data=math_train, ntree=1200, importance=TRUE)
math_pred = q5_math_rf %>% predict(newdata= math_test, type="prob")
math_test = math_test %>% mutate(.fitted_rf = math_pred[, 2])
roc = roc(math_test, x=.fitted_rf, class=pass, pos_class=1, neg_class=0)
plot(roc)
auc(roc)

#Based on the AUC value of 0.55 for this randomForest model, family structures are not important factors in comparison #to other available variables in this dataset. When only considering family related factors the AUC score of this model #decreased dramactically in comparison to Q3 and Q4 model. 
```







#===================================================================


#====================================
#======= Question 6 (1 Point) =======
# Q6. Suggest one possible application of this predictive analytics to improve education.
#     Write your answer below by using comment (#), up to 10 lines.

#====================== Write R code HERE ==========================


```{r}
#one possible application for this predictive analytics is to gain a better understanding into 
#student's performance regarding mathematics as a subject. Through these models, one can see 
#what variables may affect a student's academic performance and hinder them from achieving their 
#true potential grade. The lesson from this analysis can also helped teacher cater teaching lessons in a way
#that can reduce high risks variables that are most influential on a student's grade and may help more student 
#pass the class. 
```






#===================================================================



#===========================================================================================================
# Before submission, I recommend you restart your RStudio and run all codes.
# Please ensure that there is no error message in your code. Any code errors will get some points deduction without exception.
#===========================================================================================================


# Congratulations on your completion of the predictive analytics course!
# Wish you the best of luck!


